|

### 2024-10-21 manuscript/paper.Rmd
finished for the day.

| Trait Name  | Trait Code | Trait Description           |
|------------|-----|------------------------------------------|
|Standard Length|	`stl` |	Distance from anterior tip of premaxilla to posterior end of last vertebra (i.e., the llhypural plate)|
|Dorsal Spine Count| `mds` |	Number of dorsal spines|
|Dorsal Fin Ray Count|`mdf`|	Number of bones in the dorsal fin posterior to the third dorsal spine (i.e., soft dorsal fin rays)
|Anal Fin Ray Count|`maf`|	Number of bones in the anal fin posterior to the anal spine (i.e., soft anal fin rays)|
|Abdominal Vertebra Count|mav|	Number of vertebrae anterior to the first vertebra contacting an anal fin pterygiophore (Aguirre et al. 2014)|
|Caudal Vertebra Count|`mcv`	| Number of vertebrae posterior to and including the first vertebra contacting an anal fin pterygiophore (Aguirre et al. 2014)|
|Pterygiophore Count|`mpt`|	Number of pterygiophores anterior to but excluding the pterygiophore under the third dorsal spine|
|Pelvic Spine Length|`lps.sc`	|Length from the base of one pelvic spine at its pelvic girdle articulation to its distal tip|
|Ectocoracoid Length|`ect.sc`|	Length from the anterior to posterior tips of the shoulder girdle base|
|Pelvic Girdle Length|`tpg.sc`|	Length from the anterior to posterior tips along midline. If vestigial, the sum of longest axis along vestiges|
|Cleithrum Length|`cle.sc`|	Length from dorsal tip to ventral tip on the anterior margin of the shoulder girdle|
|Premaxilla Length|`pmx.sc`|	Length from the anterior tip of the premaxilla to the distal tip of its ascending process|
|Dorsal Spine Length|`Ds#.sc #=1/2/3`|	Length from the dorsal spine base at the pterygiophore to its distal tip|
|Pterygiophore Length|`lpt.sc`|	Distance from anterior to posterior edges of the pterygiophore immediately before ds3 (when present)|

@@

## Missing data imputation, including fossil sex
Briefly, we used multiple imputation (@little2002statistical) to impute the sex of the fossils. Let $\boldsymbol{W}$ be an $(n_{extant} + n_{fossil}) \times 1$ vector of the covariate sex of the stickleback fish and $\boldsymbol{Y}$ be an $(n_{extant} + n_{fossil}) \times K$ matrix of the $K$ phenotypes of interest.  Because the sex of the fossilized stickleback fish is unobservable, we further define $\boldsymbol{W} = (\boldsymbol{W}_{extant}^T,\boldsymbol{W}_{fossil}^T)^T$ where $\boldsymbol{W}_{extant}$ and $\boldsymbol{W}_{fossil}$ are the $n_{extant} \times 1$ and $n_{fossil} \times 1$ vectors of the observed extant sex and missing fossil sex, respectively. We imputed missing sex for the fossil data by sampling from the posterior predictive distribution $P(\boldsymbol{W}_{fossil}|\boldsymbol{W}_{extant},\boldsymbol{Y})$ using the multiple imputation by chained equations (MICE) algorithm (@MICE) with predictive mean matching, implemented in R (CITATION).  Traditionally, the choice for the number of completed dats sets is a relatively small number such as $M = 5$ or $M = 10$.  However, @ZhouReiter2010 recommend a larger number of imputed data sets if users intend on performing Bayesian analysis after imputation.  Therefore, we imputed $M = 100$ complete datasets. \textcolor{red}{PRESUMABLY this also imputes missing data in the other traits? We should probably say that explicitly. -YS}

\textcolor{red}{AKHIL'S METHODS FOR VALIDATION OF THE IMPUTATION MODEL GOES HERE}

@@

## Modeling fossil stickleback trait means through time for males and females
The goal of our analysis was to estimate fossil trait means through time, by sex, while incorporating uncertainty across imputations in the classification of sex. We used a Bayesian approach to estimate trait means in part because imputation is the process of sampling from a posterior predictive distribution of missing data, given observed data. \textcolor{red}{THIS NEEDS MORE WORK TO JUSTIFY THE BAYESIAN APPROACH, EXPLAINING IT TO MY AUDIENCE. -YS}

@@

## Missing Data Imputation
\textcolor{red}{WE NEED RESULTS SECTION DESCRIBING THE RESULTS OF AKHIL'S MULTIPLE IMPUTATION, BOTH THE TESTS OF HOW WELL THE MODEL WORKED CLASSIFYING MODEL SAMPLES, AND SUMMARY STATISTICS OF FOSSIL IMPUTATION RESULTS -YS}

>>>>

| Trait Name  | Trait Code | Trait Description           |
|------------|-----|------------------------------------------|
|Standard Length|	`stl` |	Distance from anterior tip of premaxilla to posterior end of last vertebra (i.e., the llhypural plate)|
|Dorsal Spine Count| `mds` |	Number of dorsal spines|
|Dorsal Fin Ray Count|`mdf`|	Number of bones in the dorsal fin posterior to the third dorsal spine (i.e., soft dorsal fin rays)
|Anal Fin Ray Count|`maf`|	Number of bones in the anal fin posterior to the anal spine (i.e., soft anal fin rays)|
|Abdominal Vertebra Count|`mav`|	Number of vertebrae anterior to the first vertebra contacting an anal fin pterygiophore (Aguirre et al. 2014)|
|Caudal Vertebra Count|`mcv`	| Vertebrae including and posterior to the first that contacts an anal fin pterygiophore (Aguirre et al. 2014)|
|Pterygiophore Count|`mpt`|	Number of pterygiophores anterior to but excluding the pterygiophore under the third dorsal spine|
|Pelvic Spine Length|`lps.sc`	|Length from the base of one pelvic spine at its pelvic girdle articulation to its distal tip|
|Ectocoracoid Length|`ect.sc`|	Length from the anterior to posterior tips of the shoulder girdle base|
|Pelvic Girdle Length|`tpg.sc`|	Length from the anterior to posterior tips along midline. If vestigial, the sum of longest axis along vestiges|
|Cleithrum Length|`cle.sc`|	Length from dorsal tip to ventral tip on the anterior margin of the shoulder girdle|
|Premaxilla Length|`pmx.sc`|	Length from the anterior tip of the premaxilla to the distal tip of its ascending process|
|Dorsal Spine Length|`Ds#.sc #=1/2/3`|	Length from the dorsal spine base at the pterygiophore to its distal tip|
|Pterygiophore Length|`lpt.sc`|	Distance from anterior to posterior edges of the pterygiophore immediately before ds3 (when present)|

@@

## Missing data imputation, including fossil sex
Briefly, we used multiple imputation (@little2002statistical) to impute the sex of the fossils. Let $\boldsymbol{W}$ be an $(n_{extant} + n_{fossil}) \times 1$ vector of the covariate sex of the stickleback fish and $\boldsymbol{Y}$ be an $(n_{extant} + n_{fossil}) \times K$ matrix of the $K$ phenotypes of interest.  Because the sex of the fossilized stickleback fish is unobservable, we further define $\boldsymbol{W} = (\boldsymbol{W}_{extant}^T,\boldsymbol{W}_{fossil}^T)^T$ where $\boldsymbol{W}_{extant}$ and $\boldsymbol{W}_{fossil}$ are the $n_{extant} \times 1$ and $n_{fossil} \times 1$ vectors of the observed extant sex and missing fossil sex, respectively. We imputed missing sex for the fossil data by sampling from the posterior predictive distribution $P(\boldsymbol{W}_{fossil}|\boldsymbol{W}_{extant},\boldsymbol{Y})$ using the multiple imputation by chained equations (MICE) algorithm (@MICE) with predictive mean matching, implemented in R (CITATION).  Traditionally, the choice for the number of completed dats sets is a relatively small number such as $M = 5$ or $M = 10$.  However, @ZhouReiter2010 recommend a larger number of imputed data sets if users intend on performing Bayesian analysis after imputation.  Therefore, we imputed $M = 100$ complete datasets. 
\textcolor{red}{Greg, PRESUMABLY this also imputes missing data in the other traits? We should probably say that explicitly. -YS}
\textcolor{red}{Greg, AKHIL'S METHODS FOR VALIDATION OF THE IMPUTATION MODEL GOES HERE}

@@

## Modeling fossil stickleback trait means through time for males and females
The goal of our analysis was to estimate fossil trait means through time, by sex, while incorporating uncertainty across imputations in the classification of sex. We used a Bayesian approach to estimate trait means in part because imputation is the process of sampling from a posterior predictive distribution of missing data, given observed data. \textcolor{red}{Matt, THIS NEEDS MORE WORK TO JUSTIFY THE BAYESIAN APPROACH, EXPLAINING IT TO MY AUDIENCE. -YS}

@@

## Missing Data Imputation
\textcolor{red}{Greg, WE NEED RESULTS SECTION DESCRIBING THE RESULTS OF AKHIL'S MULTIPLE IMPUTATION, BOTH THE TESTS OF HOW WELL THE MODEL WORKED CLASSIFYING MODEL SAMPLES, AND SUMMARY STATISTICS OF FOSSIL IMPUTATION RESULTS -YS}
